---
title: "a) Ecuador analysisn - Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a) Ecuador analysisn - Overview}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(avesdemazan)
```


Files received from Emily Morrison via email.  EM indicated  
"You should only need
Ecuador_Analysis_DataPrep.Rmd
Ecuador_Analysis.Rmd
Net_hours.csv
Banding Sheet 2.csv
I included my version of intermediate data prep in case things are broken you can see where things were supposed to end up."


Scripts provided:

* MAIN: Ecuador_Analysis_DataPrep.Rmd
* MAIN: Ecuador_Analysis.Rmd

Data provided:

* MAIN: Net_hours.csv
* MAIN: Banding Sheet 2.csv
* ecuador_working.csv
* ecuador_1.csv, ecuador_2, ecuador_3, _4, _5



# Files and dataframes used

Mist net data description: ecuador_working.csv
used to make ecuador_id dataframe
(probably included untagged individuals)
size: 4708 x 23

Point count data: ecuador_pc (all commented out)

regression analysis uses "ecuador" dataframe from the ecuador_5.csv file
size: 6860 x  11


