---
title: "f) Mist net data prep - NOT RUN"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{f) Mist net data prep - NOT RUN}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = F
)
```

```{r setup}
library(avesdemazan)
```

This is a version of the data cleaning/prep script used by Emily Scott to prepare the data.  As of June 4 2021 it has not been checked or integrated into this compendium and is assumed to be correct.

The original file name was Ecuador_Analysis_DataPrep.Rmd

## Load packages
```{r, warning = FALSE, message = FALSE}
library(lme4)
library(arm)
library(tidyverse)
library(reshape2)
library(gridExtra)
library(rgl)
library(stringi)
```

## Load data
```{r, warning = FALSE, message = FALSE}
# Location
#loc <- "C:/Users/emily/OneDrive/Desktop/Ecuador"

# Load uncleaned data set 
#ecuador <- read.csv(file = paste0(loc, "/Banding Sheet 2.csv"))

# Load species trait data
#spp_trait <- read.csv(file = paste0(loc, "/Ecuador_Species_List.csv"))

# Load net hours
#net_hours <- read.csv(file = paste0(loc, "/Net_Hours.csv"))
```

## Preliminary look at data
```{r}
# Column names
names(ecuador)

# Rename first column to "RU" from ?..RU 
colnames(ecuador)[1] <- "RU"

# Check column names again
names(ecuador)
```

## RU: An identity number for each record of the base. A unique record. 
```{r}
# Make sure there are no duplicates or missing entries
summary(ecuador$RU)
which(ecuador$RU == "")
which(is.na(ecuador$RU) == TRUE)
```

## Band number: The unique number of each band to identify an individual.
```{r}
# Summary
summary(ecuador$Band.Number)

# Find all "99999" entries
i.band.number <- which(ecuador$Band.Number == "99999"|
                       ecuador$Band.Number == "SIN ANILLO")

# Store those cells in table for inspection
prep.band.number <- ecuador[i.band.number,]
prep.band.number

# Fill 99999 cells with "NA"
ecuador$Band.Number[i.band.number] <- NA

# Double-check for empty cells
which(ecuador$Band.Number == "")

# Fill empty cell
ecuador$Band.Number[which(ecuador$Band.Number == "")] <- NA

# Summarize and store results
# Confirm empty cells have been filled with "NA" 
summary.band.number <- summary(ecuador$Band.Number)
summary.band.number
```

## Code: Band code, N: new, R: recapture, D: destroyed, L: Lost, U: unbanded.
```{r}
# Summary
summary(ecuador$Code)

# Find all entries that contain entries (N, R, D, L, U)
i.code <- which(ecuador$Code %in% c("N","R","D","L","U"))

# Create table that contains all non-standard entries
prep.code <- ecuador[-(i.code),]
prep.code

# Fill empty cells with "NA"
ecuador$Code[which(ecuador$Code == "")] <- NA

# Summarize and store results
# Confirm that empty cells have been filled with "NA" 
summary.code <- summary(ecuador$Code)
summary.code

# Double-check for empty cells
which(ecuador$Code == "")
```

## Species name: the scientific name of the species.
```{r}
# Summary
summary(ecuador$Specie.Name)

# Convert to character
ecuador$Specie.Name <- as.character(ecuador$Specie.Name)

# Find all entries that are either empty or contain "99999"
i.spp.name <- which(ecuador$Specie.Name == "" | 
                    ecuador$Specie.Name == "99999" | 
                    ecuador$Specie.Name == "9" |
                    ecuador$Specie.Name == "99"|
                    ecuador$Specie.Name == "U")

# Create table that contains all cells missing species information
prep.spp.name <- ecuador[i.spp.name,]
prep.spp.name

# Overwrite cells missing species information with "NA"
ecuador$Specie.Name[i.spp.name] <- NA

# Summarize and store results
# Confirm that cells missing species information have been filled with "NA" (expect 94)
summary.spp.name <- summary(ecuador$Specie.Name)
summary.spp.name

# Double-check for empty cells
which(ecuador$Specie.Name == "")

# Email from Steve Dec 06, 2017
# There is a tapaculo that was split into two species. 
# It looks like we have both in our data base but they should ALL be S. latrans 
# (S. unicolor is the old species name before the split and now represents a different species). 
# Change the unicolor to latrans so we only have one.
# No instances of S. unicolor
which(ecuador$Specie.Name == "SCYTALOPUS UNICOLOR")

# Fix POROARIA CORONATA --> PAROARIA CORONATA
ecuador[which(ecuador$Specie.Name == "POROARIA CORONATA"),"Specie.Name"] <- "PAROARIA CORONATA"
```

## Species code: the first two letters of the genus and the first two letters of the species.
```{r}
# Summary
summary(ecuador$Specie.Code)

# Change Specie.Code to character
ecuador$Specie.Code <- as.character(ecuador$Specie.Code)

# Find all entries that are either empty or contain "9999"
i.spp.code <- which(ecuador$Specie.Code == "" | 
                    ecuador$Specie.Code == "9999" | 
                    ecuador$Specie.Code == "9"|
                    ecuador$Specie.Code == "U")

# Create table that contains all cells missing species code information
prep.spp.code <- ecuador[i.spp.code,]
prep.spp.code

# Overwrite cells missing species code information with "NA"
ecuador$Specie.Code[i.spp.code] <- NA

# Summarize and store results
# Confirm that cells missing species code information have been filled with "NA" 
summary.spp.code <- summary(ecuador$Specie.Code)
summary.spp.code

# Double-check for empty cells
which(ecuador$Specie.Code == "")

# Fix POCO to PACO
# POROARIA CORONATA --> PAROARIA CORONATA
i.POCO <- which(ecuador$Specie.Code == "POCO")
ecuador[i.POCO,"Specie.Code"] <- "PACO"

# Fix CHMU to CHVE
i.CHMU <- which(ecuador$Specie.Code == "CHMU")
ecuador[i.CHMU,"Specie.Code"] <- "CHVE"
```

## Date: The date that the bird was captured or recaptured.
```{r}
# Look at date column
summary(ecuador$Date)

# Format: dd-mm-yyyy
ecuador$date2 <- as.Date(ecuador$Date, format = "%d-%b-%y")

## Create columns for year, month, day
library(lubridate)
ecuador$year  <- year(ecuador$date2)
ecuador$month <- month(ecuador$date2)
ecuador$day   <- day(ecuador$date2)
summary(ecuador$date2)

# Fix Segunda 13 date 
i.date <- which(ecuador$date2 == "2012-10-11" & ecuador$Session == "SEGUNDA 13")
ecuador[i.date,"date2"] <- as.Date("11-Oct-13", format = "%d-%b-%y")
ecuador[which(ecuador$Session == "SEGUNDA 13"),]
```

## Location: There are two study locations: a) LLAVIUCU and b) MAZAN, and two sites per locations. 
## In location a) the sites are a.1) LLAV and a.2) SANA 
## In location b) the sites are b.1) MASE and B.2) MAIN
```{r}
# Summary
summary(ecuador$Location)

# Find those cells
i.location <- which(ecuador$Location == "")

# Create table that contains these entries
prep.location <- ecuador[i.location,]
prep.location

# Fill empty cells with "NA"
ecuador$Location[which(ecuador$Location == "")] <- NA

# Summarize and store results
# Confirm that empty cells have been filled with "NA" 
summary.location <- summary(ecuador$Location)
summary.location

# Double-check for empty cells
which(ecuador$Location == "")
```

## Location2: Create column containing Llaviucu and Mazan (i.e. group LLAV & SANA, and group MASE & MAIN)
```{r}
# Index locations
i.llaviucu <- which(ecuador$Location == "LLAV" | ecuador$Location == "SANA")
i.mazan    <- which(ecuador$Location == "MASE" | ecuador$Location == "MAIN")

# Create column to store location
ecuador[i.llaviucu,"location2"] <- "Llaviucu"
ecuador[i.mazan,"location2"] <- "Mazan"

# Change entries to be factors
ecuador$location2 <- as.factor(ecuador$location2)

# Look at location summary
summary(ecuador$location2)
```

# Habitat type
```{r}
# Create empty habitat column
ecuador$Habitat <- NA

# MASE is primary forest
ecuador[which(ecuador$Location == "MASE"),]$Habitat <- "forest.prim"

# MAIN is introduced forest
ecuador[which(ecuador$Location == "MAIN"),]$Habitat <- "forest.intro"

# LLAV is secondary forest
ecuador[which(ecuador$Location == "LLAV"),]$Habitat <- "forest.sec"
```

# Session
```{r}
# Create new column that doesn't include year (i.e. PRIMERA 16 --> PRIMERA; SEGUNDA 16 --> SEGUNDA; etc)
ecuador$session2 <- substr(ecuador$Session, 0, 7)

# Manually fix error (Late September sampling categorized as primera should be tercera)
ecuador[which(ecuador$Session == "PRIMERA 16" & ecuador$month == "9"),"session2"] <- "TERCERA"

# Extra session data entered as EXTRA 2016
# Re-enter as SEGUNDA so it gets grouped with appropriate data
ecuador[which(ecuador$Session == "EXTRA 2016"),"session2"] <- "SEGUNDA"
```

# Merge species traits
```{r}
ecuador$Specie.Code   <- as.character(ecuador$Specie.Code)
spp_trait$Specie.Code <- as.character(spp_trait$Specie.Code)
ecuador <- left_join(ecuador, spp_trait, by = "Specie.Code")
```

# Omit species
```{r}
# Steve said to omit ACVE, NOMU
# From Dec 6, 2017 email: There are a few species to omit (things that generally fly over and so are excluded from the analyses)
 
# Index
i.ACVE <- which(ecuador$Specie.Code == "ACVE")
ecuador <- ecuador[-i.ACVE,]
i.NOMU <- which(ecuador$Specie.Code == "NOMU")
ecuador <- ecuador[-i.NOMU,]
```

# Create dataset that includes everything (Dataset 1)
```{r, message = FALSE, warning = FALSE}
write.csv(ecuador, paste0(loc, "/ecuador_1.csv"))
```

# Create dataset that excludes SANA (Dataset 2)
```{r, message = FALSE, warning = FALSE}
# Index SANA
i.SANA <- which(ecuador$Location == "SANA")
ecuador2 <- ecuador[-i.SANA,]
write.csv(ecuador2, paste0(loc, "/ecuador_2.csv"))
```

# Create dataset that excludes SANA, SEGUNDA 13, TERCERA 13 (Jan), PRIMERA 16, March 2017 data (Dataset 3)
```{r, message = FALSE, warning = FALSE}
i.segunda13 <- which(ecuador2$Session == "SEGUNDA 13")
i.tercera13 <- which(ecuador2$Session == "TERCERA 13" & ecuador2$month == 1)
i.primera16 <- which(ecuador2$Session == "PRIMERA 16")
i.2017      <- which(ecuador2$year == "2017")
i.drop <- c(i.segunda13, i.tercera13, i.primera16, i.2017)
ecuador3 <- ecuador2[-i.drop,]
write.csv(ecuador3, paste0(loc, "/ecuador_3.csv"))
```

# Create dataset that exluces SEGUNDA 13, TERCERA 13, PRIMERA 16, PRIMERA 17 (Dataset 4)
```{r, message = FALSE, warning = FALSE}
i.segunda13 <- which(ecuador2$Session == "SEGUNDA 13")
i.tercera13 <- which(ecuador2$Session == "TERCERA 13")
i.primera16 <- which(ecuador2$Session == "PRIMERA 16")
i.2017      <- which(ecuador2$year == "2017")
i.drop <- c(i.segunda13, i.tercera13, i.primera16, i.2017)
ecuador4 <- ecuador2[-i.drop,]
write.csv(ecuador4, paste0(loc, "/ecuador_4.csv"))
```

# Create working dataset
# Use dataset 3
# Retain the relevant columns
```{r, message = FALSE, warning = FALSE}
# Create working dataset
ecuador_working <- ecuador3[,c("Band.Number","Band.Size","Specie.Name","Specie.Code",
                               "Date","Location","Session","date2","year",
                               "month","day","location2","Habitat","session2",
                               "Body_Size", "Diet", "Foraging_Strata", "Primary_Habitat",
                               "Habitat_Breadth","Rarity_Range_Size", "Rarity_Local_Abund", 
                               "Rarity_Neotrop_Abund")]
# Remove rows with missing species code
i.drop <- which(is.na(ecuador_working$Specie.Code) == TRUE)
ecuador_working <- ecuador_working[-i.drop,]
write.csv(ecuador_working, paste0(loc, "/ecuador_working.csv"))
```

# Create dataframe in long format that includes net hours
```{r}
# Prepare net hours dataframe
net_hours$Date2 <- as.Date(net_hours$Date, format = "%m/%d/%Y")
net_hours$year <- year(net_hours$Date2)
net_hours$session2 <- substr(net_hours$Session, 0, 7)

# Remove session == EXTRA 2016 and location == SANA (not used in analysis)
net_hours <- net_hours[-which(net_hours$Session == "EXTRA 2016"),]
net_hours <- net_hours[-which(net_hours$Location == "SANA"),]

# Remove 2006 and 2007 because some data is missing 
# Values added by hand subsequently
net_hours <- net_hours[-which(net_hours$year == 2006 | net_hours$year == 2007),]

# Create variable to merge net hours with capture data
net_hours$groupby <- paste(net_hours$session2, net_hours$year, net_hours$Location, sep = ".")
net_hours2 <- group_by(net_hours, groupby) %>%
  summarize(tot_net_hours = sum(Total.Net.Hours))
net_hours2$groupby <- as.character(net_hours2$groupby)

# Recreate year, location, session columns
net_hours2$session2 <- substr(net_hours2$groupby, 1, 7)
net_hours2$year <- substr(net_hours2$groupby, 9, 12)
net_hours2$Location <- substr(net_hours2$groupby, 14, 17)

# Add missing data
net_hours_missing <- data.frame(year = c(rep(2006, 9), rep(2007, 9)),
                                Location = c(rep("LLAV", 3), rep("MAIN", 3), rep("MASE", 3),
                                             rep("LLAV", 3), rep("MAIN", 3), rep("MASE", 3)),
                                session2 = c(rep(c("PRIMERA", "SEGUNDA", "TERCERA"), 6)),
                                tot_net_hours = c(256.3, 291.6, 282.0, 
                                                  326.6, 245.2, 160.0, 
                                                  300.9, 288.4, 220.0,
                                                  311.0, 256.6, 320.0,
                                                  237.5, 233.4, 251.6,
                                                  281.6, 293.4, 300.0))

net_hours_missing$groupby <- paste(net_hours_missing$session2, net_hours_missing$year, net_hours_missing$Location, sep = ".")

# Bind existing data with manually entered data
net_hours <- rbind(net_hours2, net_hours_missing)
```

```{r}
# Remove capture data entries from EXTRA 2016 
i.drop <- which(ecuador_working$Session == "EXTRA 2016")
e1 <- ecuador_working[-i.drop,]

# Reshape data to long format
e2 <- dcast(e1, formula = Specie.Code + Location + session2 + year ~ . , value.var = "Band.Number", length)
colnames(e2)[5] <- "N"

# Add zeros
merge <- expand.grid(Specie.Code = unique(e2$Specie.Code),
                     year        = unique(e2$year),
                     session2    = unique(e2$session2),
                     Location    = unique(e2$Location))
merge$merge <- paste(merge$Specie.Code, merge$Location, merge$session2, sep = ".")
e2$merge    <- paste(e2$Specie.Code,    e2$Location,    e2$session2,    sep = ".")
e3 <- merge(merge, e2, by = c("merge", "year"), all = TRUE)
i.zero <- which(is.na(e3$N))
e3[i.zero,"N"] <- 0
e3$i <- 1:dim(e3)[1]
e3 <- e3[,c("Specie.Code.x","year","session2.x","Location.x","N","i")]
names(e3) <- c("Specie.Code", "year", "session", "Location", "N", "i")
# Create variable to merge capture data with net hours
e3$groupby <- paste(e3$session, e3$year, e3$Location, sep = ".")
e3$groupby <- as.character(e3$groupby)
# Merge capture data with net hours
e4 <- left_join(e3, net_hours, by = "groupby")
# Add caps per 1000 net hours
e4$caps_per_1K_nethours <- with(e4, N/tot_net_hours * 1000)
# Aside: reshape to wide format
e5 <- dcast(e1, formula = Specie.Code + Location + session2 ~ year, value.var = "year", length)
# Define function to modify entries to presence/absence
fx01 <- function(x)
{
  ifelse(x > 0, 1, 0)
}
# Convert to presence/absence 
e5[ , -c(1:3)] <- apply(e5[ -c(1:3)], 2, fx01)
# Sum rows to calculate total number of years captured
e5$tot.yrs <- rowSums(e5[ , -c(1:3)])
# Remove rows of data associated with annual captures
e5 <- e5[ , -c(4:14)]
# Add total years captured to working dataframe
e4$groupby2 <- paste(e4$Specie.Code, e4$session,  e4$Location.x, sep = ".")
e5$groupby2 <- paste(e5$Specie.Code, e5$session2, e5$Location, sep = ".")
e6 <- left_join(e4, e5, by = "groupby2")
# Fill in zeros for species/session/location combos that were never captured
e6[which(is.na(e6$tot.yrs) == TRUE),"tot.yrs"] <- 0
# Fix column names
e6 <- e6[,c("Specie.Code.x", "year.x", "session", "Location.x", "N", "i",
            "tot_net_hours", "caps_per_1K_nethours", "tot.yrs")]
colnames(e6) <- c("Specie.Code", "year", "session", "Location", "N", "i",
                  "tot_net_hours", "caps_per_1K_nethours", "tot.yrs")
# Add centered year
e6$year_cent <- scale(e6$year, scale = FALSE)
e6$year_cent <- as.numeric(e6$year_cent)
# Remove NAs 
e7 <- e6[-which(is.na(e6$tot_net_hours) == TRUE),]
```

# Save new working dataset
```{r, warning = FALSE, message = FALSE}
write.csv(e7, paste0(loc,"/ecuador_5.csv"))
```
