---
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r}
fm1 <- glmmTMB(count~mined+(1|spp),
                  ziformula=~mined,
                  data=Salamanders,
                  family=nbinom1)


## single parametric bootstrap step: refit with data simulated from original model
fm1R <- refit(fm1, simulate(fm1)[[1]])
   
## the bootMer function from lme4 provides a wrapper for doing multiple refits
##   with a specified summary function
fixef(fm1R)
fixef(fm1R)$zi

b1 <- lme4::bootMer(fm1, FUN=function(x) fixef(x)$zi, nsim=5, .progress="txt")

   if (requireNamespace("boot")) {
      boot:boot.ci(b1,type="perc")
    }
```



Simulate from model
```{r}
simulate(best_randslopes)
```


Get output as vector
```{r}
simulate(best_randslopes)[[1]]
```


Refit
```{r}
best_randslopes_refit <- refit(best_randslopes, simulate(best_randslopes)[[1]])

fixef(best_randslopes_refit)

```

bootstrap fixedeffects 

boot.ci has 
index = 1:min(2,length(boot.out$t0))

use.u	
logical, indicating whether the spherical random effects should be simulated / bootstrapped as well. If TRUE, they are not changed, and all inference is conditional on these values. If FALSE, new normal deviates are drawn (see Details).

re.form	
formula, NA (equivalent to use.u=FALSE), or NULL (equivalent to use.u=TRUE): alternative to use.u for specifying which random effects to incorporate. See simulate.merMod for details.


```{r}
fixef(best_randslopes_refit)$cond[1]

best_randslopes <- fit_pois_corr_FULL_RF

boot_fnxn_fixef <- function(x)
  {
  fixef(x)$cond
  }

x1 <- lme4::bootMer(best_randslopes, 
                    FUN = boot_fnxn_fixef,
                    nsim=3, .progress="txt",
                    use.u = FALSE, 
                    re.form = NA)
boot.ci(x1, type = "norm")  #
boot.ci(x1, type = "basic") # warning "extreme order statistics used as endpoints"
boot.ci(x1, type = "perc")  # warning "extreme order statistics used as endpoints"
# boot.ci(x1, type = "stud")  # error
# boot.ci(x1, type = "bca")   # error
``` 


```{r}
ranef(best_randslopes)$cond$`Specie.Code:Location`$time_cts
```

## Functions

```{r}
boot_fnxn_fixefs <- function(x)
  {
  fixef(x)$cond
  }

boot_fnxn_randefs <- function(x) 
  {
  ranef(x)$cond$`Specie.Code:Location`$time_cts
}

boot_fnxn_randslopes_combo <- function(x) 
  {
  fixedef <- fixef(x)$cond["time_cts"]
  randslope <- ranef(x)$cond$`Specie.Code:Location`$time_cts
  
  combo <- fixedef + randslope
}
```


## Bootstrap fixeffect
```{r}
fixef(best_randslopes)$cond["time_cts"]
```


```{r}
boot_fixef <- lme4::bootMer(best_randslopes, 
                    FUN=boot_fnxn_fixefs,
                    nsim=500, .progress="txt",
                    use.u = FALSE, 
                    re.form = NA)


```


## Bootstrap randomeffect

```{r}
ranef(best_randslopes)$cond$`Specie.Code:Location`$time_cts
```

```{r}
boot_randeff <- lme4::bootMer(best_randslopes, 
                    FUN=boot_fnxn_randefs,
                    nsim=500, .progress="txt",
                    use.u = FALSE, 
                    re.form = NA)
```



```{r}
boot_randslopes_combo <- lme4::bootMer(best_randslopes, 
                    FUN=boot_fnxn_randslopes_combo,
                    nsim=500, .progress="txt",
                    use.u = FALSE, 
                    re.form = NA)
```

```{r}
x1$t0
x1$t
```

```{r}
boot.ci(boot_randslope, type = "norm", index = 1)  #
boot.ci(boot_randslope, type = "norm", index = 2)  #
boot.ci(x1, type = "norm", index = 3)  #

boot.ci(x1, type = "basic") # warning "extreme order statistics used as endpoints"
boot.ci(x1, type = "perc")  #
```

```{r}
 head(as.data.frame(x1))
```


```{r}
B_best_randslopies <- ranef(best_randslopes)$cond$`Specie.Code:Location`$time_cts
B_labels <- rownames(ranef(best_randslopes)$cond$`Specie.Code:Location`)
B.n <- length(B_best_randslopies)

```

```{r}
get_CI <- function(x,w) {
   b1 <- boot.ci(x,index=w, type = "norm")
   ## extract info for all CI types
   tab <- t(sapply(b1[-(1:3)],function(x) tail(c(x),2)))
   ## combine with metadata: CI method, index
   tab <- cbind(w,rownames(tab),as.data.frame(tab))
   colnames(tab) <- c("i","meth","lwr","upr")
   tab
}
## do it for both parameters
x <- do.call(rbind,lapply(1:B.n,get_CI,x=boot_randslopes_combo))

which(sign(x$lwr) == sign(x$upr))


build_CI_tab <- function(boot_out, betas, B_labels){
  CIs <- do.call(rbind,lapply(1:B.n,get_CI,x=boot_out))
  beta.sign <- ifelse(betas < 0, "-", "+")
  sig <- ifelse(sign(CIs$lwr) == sign(CIs$upr),"*","")
  CI_tab <- cbind(B_labels,
                  betas,
                  CIs[3:4],
                  beta.sign,
                  sig,CIs[1:2])
  rownames(CI_tab) <- NULL
  CI_tab$method <- gsub("normal","norm",CI_tab$method)
  return(CI_tab)
}

build_CI_tab(boot_randslopes_combo, B_best_randslopies, B_labels)



```


```{r}
x1$t0
```

