---
title: "Untitled"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**NOTE: these random slopes are deviations from the OVERALL slope parameter of the model!**


## Process random slopes


Extract SD / var using TMB::sdreport()

"After optimization of an AD model, sdreport() is used to calculate standard deviations of all model parameters"

* `getJointPrecision	= Optional`. Return full joint precision matrix of random effects and parameters?
  + From `sdreport` doc:"The full joint covariance is not returned by default, because it may require large amounts of memory. It may be obtained by specifying `getJointPrecision=TRUE`, ..."
* `bias.correct`	= logical indicating if bias correction should be applied

Data


```{r}
# This is the best model structure
data(fit_pois_corr_FULL_RF)

# This is a reduced model structure
data(fit_pois_corr_only_sppLoc)

# mis-specified models, for comparison
data(fit_nb2_corr_misspec)  # nb mis-spec
data(fit_pois_corr_misspec) # pois mis-spec


```



```{r}
str_ranef <- function(mod){
  str(ranef(mod)[["cond"]],max.level = 1, vec.leng = 1, digits.d =2,
      give.length = F, give.head = F)
}
```


```{r}
str_ranef(fit_pois_corr_FULL_RF)
str_ranef(fit_pois_corr_only_sppLoc)# NEEDS UPDATE
str_ranef(fit_nb2_corr_misspec)
str_ranef(fit_pois_corr_misspec)

diagnose(fit_pois_corr_FULL_RF)
diagnose(fit_pois_corr_only_sppLoc)  # issues
diagnose(fit_nb2_corr_misspec) 
diagnose(fit_pois_corr_misspec) 
```

```{r}
summary(fit_pois_corr_FULL_RF)$varcor
summary(fit_pois_corr_only_sppLoc)$varcor
summary(fit_nb2_corr_misspec)$varcor
summary(fit_pois_corr_misspec)$varcor
```


```{r}
summary(fit_pois_corr_FULL_RF)$coefficients$cond
summary(fit_pois_corr_FULL_RF)$varcor
```


Get CIs

```{r}
# add metadata
data("Ecuador_Species_List")


str(ranef(fit_nb2_corr_misspec),1)

str(ranef(fit_nb2_corr_misspec)[["cond"]],1)

slopes_FULL_RF <- glmmTMB_randslope_ci(fit_pois_corr_FULL_RF, 
                                           randeff = "Specie.Code:Location")
slopes_only_sppLoc <- glmmTMB_randslope_ci(fit_pois_corr_only_sppLoc, 
                                           randeff = "Specie.Code:Location")

slopes_nb2_misspec <- glmmTMB_randslope_ci(fit_nb2_corr_misspec, 
                                           randeff = "Specie.Code:Location")
slopes_pois_misspec <- glmmTMB_randslope_ci(fit_pois_corr_misspec, 
                                           randeff = "Specie.Code:Location")




slopes_FULL_RF_final <- add_mazan_metadata(spp_list = Ecuador_Species_List,
                                   slopes = slopes_FULL_RF)
slopes_only_sppLoc_final <- add_mazan_metadata(spp_list = Ecuador_Species_List,
                                   slopes = slopes_only_sppLoc)

slopes_nb2_misspec_final <- add_mazan_metadata(spp_list = Ecuador_Species_List,
                                   slopes = slopes_nb2_misspec)

    
slopes_pois_misspec_final <- add_mazan_metadata(spp_list = Ecuador_Species_List,
                                   slopes = slopes_pois_misspec)

```

```{r}
summary(slopes_FULL_RF_final$Sig)
summary(slopes_only_sppLoc_final$Sig)
summary(slopes_nb2_misspec_final$Sig)
summary(slopes_pois_misspec_final$Sig)
```



```{r}
slopes_FULL_RF_final[1,c("time_lb2","time_cts_tot","time_ub2")]
par(mfrow = c(1,1), mar = c(2,2,0,0))
plot(slopes_FULL_RF_final$time_cts_tot , slopes_only_sppLoc_final$time_cts_tot ,
     xlim = c(-.52185,.5),
     ylim = c(-.52185,.5))
segments(x0 = slopes_FULL_RF_final$time_cts_tot,
         x1 = slopes_FULL_RF_final$time_cts_tot,
         y0 = slopes_FULL_RF_final$time_lb2,
         y1 = slopes_FULL_RF_final$time_ub2)
segments(y0 = slopes_only_sppLoc_final$time_cts_tot,
         y1 = slopes_only_sppLoc_final$time_cts_tot,
         x0 = slopes_only_sppLoc_final$time_lb2,
         x1 = slopes_only_sppLoc_final$time_ub2)
abline(a = 0, b = 1)
abline(v = 0)
abline(h = 0)

plot(slopes_FULL_RF_final$time_cts_tot , slopes_nb2_misspec_final$time_cts_tot )
abline(a = 0, b = 1)
abline(v = 0)
abline(h = 0)

plot(slopes_FULL_RF_final$time_cts_tot , slopes_pois_misspec_final$time_cts_tot ,
     xlim = c(-.52185,0.0543),
     ylim = c(-.52185,0.0543))
abline(a = 0, b = 1, lwd = 3, col =2)
abline(v = 0, b = 1, lwd = 3, col =2)
abline(h = 0, b = 1, lwd = 3, col =2)
segments(x0 = slopes_FULL_RF_final$time_cts_tot,
         x1 = slopes_FULL_RF_final$time_cts_tot,
         y0 = slopes_FULL_RF_final$time_lb2,
         y1 = slopes_FULL_RF_final$time_ub2, col =4)
segments(y0 = slopes_pois_misspec_final$time_cts_tot,
         y1 = slopes_pois_misspec_final$time_cts_tot,
         x0 = slopes_pois_misspec_final$time_lb2,
         x1 = slopes_pois_misspec_final$time_ub2)

```






data()

# This is a reduced model structure
data()

# mis-specified models, for comparison
data()  # nb mis-spec
data() # pois mis-spec

```{r}
fit_pois_corr_FULL_RF_boot <- lme4::bootMer(fit_pois_corr_FULL_RF, 
                    FUN=boot_fnxn_randslopes_combo,
                    nsim=1000, .progress="txt",
                    use.u = FALSE, 
                    re.form = NA)

fit_pois_corr_only_sppLoc_boot <- lme4::bootMer(fit_pois_corr_only_sppLoc, 
                    FUN=boot_fnxn_randslopes_combo,
                    nsim=500, .progress="txt",
                    use.u = FALSE, 
                    re.form = NA)

fit_nb2_corr_misspec_boot <- lme4::bootMer(fit_nb2_corr_misspec, 
                    FUN=boot_fnxn_randslopes_combo,
                    nsim=2000, .progress="txt",
                    use.u = FALSE, 
                    re.form = NA)

fit_pois_corr_misspec_boot <- lme4::bootMer(fit_pois_corr_misspec, 
                    FUN=boot_fnxn_randslopes_combo,
                    nsim=500, .progress="txt",
                    use.u = FALSE, 
                    re.form = NA)

```

```{r}
hist(fit_nb2_corr_misspec_boot$t[,1])
x <- rep(NA, length(fit_nb2_corr_misspec_boot$t[,1]))
for(i in 1: length(x))
{
  x[i] <- mean(fit_nb2_corr_misspec_boot$t[1:i,1])
}
plot(x, xlim )
```


```{r}

mod_working <- fit_nb2_corr_misspec
ci_out_working <- fit_nb2_corr_misspec_boot
boot.ci(ci_out_working, type = "norm", index = 1) 

summary_out <- summary(ci_out_working)
B_best_randslopes <- ranef(mod_working)$cond$`Specie.Code:Location`$time_cts
B_labels <- rownames(ranef(mod_working)$cond$`Specie.Code:Location`)
B.n <- length(B_best_randslopes)

fixef_beta <- fixef(mod_working)$cond[["time_cts"]]

boot_table1 <- data.frame(B_labels,
      B_best_randslopes,
      fixef_beta,
      Beta_sum  = B_best_randslopes + fixef_beta,
      original = summary_out$original,
      bootBias = summary_out$bootBias,
      bootSE = summary_out$bootSE,
     bootMed =  summary_out$bootMed)
round(boot_table1[,-1],4)



ci_table <- data.frame(norm.lwr = rep(NA,  dim(boot_table1)[1]),
                       norm.upr = NA, 
                       basic.lwr = NA,
                       basic.upr = NA,
                       perc.lwr = NA,
                       perc.upr = NA,
                      norm.sig = NA,
                     basic.sig = NA,
                     perc.sig = NA)

for(i in 1:dim(ci_table)[1]){
  bootci.out <- boot.ci(fit_pois_corr_FULL_RF_boot, 
                        index = i,
                        type = c("norm","basic","perc"))  #
x <- bootci.out$normal[2:3]
x.sig <- ifelse(sign(x[1])==sign(x[2]),"*","")
y <- bootci.out$basic[4:5]
y.sig <- ifelse(sign(y[1])==sign(y[2]),"*","")
z <- bootci.out$percent[4:5]
z.sig <- ifelse(sign(z[1])==sign(z[2]),"*","")
ci_table[i,1:6 ] <- c(x,y,z)
ci_table[i,-c(1:6)] <- c(x.sig,y.sig,z.sig)
}


slopes_nb2_misspec_final[48:49,c("id","time_cts","time_cts_tot","time_lb2","time_ub2","Sig", "Species_SciName")]
boot_table1[48:49,]
ci_table[48:49,]
```

