---
title: "Untitled"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Data and functions
library(avesdemazan)

# Model fitting
library(lme4)       # glmer() for glmm
library(glmmTMB)    # glmm with AR-1  
# library(blme)       # bglmer (used?)

# Model evaluation
# library(afex)       # functionall_fit()
# library(multcomp)   
# library(emmeans)
library(bbmle)

data("ecuador")
```

```{r}
# NOTE: additional random effects will make it MUCH more difficult to calculate
# SE!!!!!!!!!!!!!!!!!!!!!!
form_glmm_ar1_FULL_RF <- formula(N ~ 1 +          # Intercept
                      Location +                  # Location FE; 3 level factor
                      # session +                 # session FE
                      # Location*session  +       # Location*session  FE
                      time_cts +                  # Continuous time variable
                      (1|Specie.Code) +           # species RE
                      
                      # session effects    
                      ## 1a)species:session; 
                      ###  OVERFITTING?  
                      ### this would have to due to biological factors like
                      ### altitudinal migration or other behaviors
                      ### that are consistent 
                      ### WITHIN a species ACRROSS the sites  
                      (1|Specie.Code:session) +   
                      
                      ### 1b)spp:sess:loc NEW
                      ### OVERFITTING?
                      ### location-specific behaviors
                      (1|Specie.Code:session:Location) +
                      
                      ### 2a) year:sess  
                      ### corr within the same sampling period    
                      (1|year:session) +          # 
                      
                      ### 2b) year:sess:loc NEW    
                      (1|year:session:Location) +          #   
                        
                      # not necessary b/c dropped in rand slope statement?
                      (1|Specie.Code:Location) +  # Species:location intercept 
                     
                       ar1(as.ordered(time_cts) + 0|Specie.Code:Location) +
                      
                      (time_cts + 
                       0|Specie.Code:Location) +   # Species-location RS
                      
                      offset(log(tot_net_hours))   # offset for effort
                      )
```




```{r}
fam_poisson <- "poisson"
fit_pois_corr_FULL_RF <- glmmTMB(form_glmm_ar1_FULL_RF,
                        family = fam_poisson,
                        data = ecuador)
diagnose(fit_pois_corr_FULL_RF)

fit_pois_corr_orig <- update( fit_pois_corr_FULL_RF, . ~ .
                              - (1|Specie.Code:session:Location) - (1|year:session:Location))

fit_pois_corr_no_sppSess <- update(fit_pois_corr_FULL_RF, 
                                   . ~ . -(1|Specie.Code:session))
fit_pois_corr_no_sppSessLoc <- update(fit_pois_corr_FULL_RF, 
                                   . ~ . -(1|Specie.Code:session:Location))

fit_pois_corr_no_yrSess <- update(fit_pois_corr_FULL_RF, 
                                   . ~ . -(1|year:session))

fit_pois_corr_no_yrSessLoc <- update(fit_pois_corr_FULL_RF, 
                                   . ~ . -(1|year:session:Location))


fit_pois_drop_randslope <- update(fit_pois_corr_FULL_RF,
                                  . ~ . -(time_cts + 0|Specie.Code:Location))

fit_pois_drop_randslope_from_orig <- update(fit_pois_corr_FULL_RF,
                                  . ~ . -(time_cts + 0|Specie.Code:Location)- (1|Specie.Code:session:Location) - (1|year:session:Location))

bbmle::AICtab(fit_pois_corr_FULL_RF,
              fit_pois_corr_orig,
              fit_pois_corr_no_sppSess,
              fit_pois_corr_no_sppSessLoc,
              
              fit_pois_corr_no_yrSess,
              fit_pois_corr_no_yrSessLoc,
              
              fit_pois_drop_randslope,
              fit_pois_drop_randslope_from_orig)
```



```{r}
cbind(new = fixef(fit_pois_corr_FULL_RF)$cond,
      orig = fixef(fit_pois_corr_orig)$cond)
```


Conditional model:
 Groups                       Name                                  Variance Std.Dev.
 Specie.Code                  (Intercept)                           0.200537 0.44781 
 Specie.Code.session          (Intercept)                           0.110581 0.33254 
 Specie.Code.session.Location (Intercept)                           0.044341 0.21057 
 year.session                 (Intercept)                           0.015569 0.12478 
 year.session.Location        (Intercept)                           0.032118 0.17922 
 Specie.Code.Location         (Intercept)                           0.240568 0.49048 
 Specie.Code.Location.1       as.ordered(time_cts)0.333333333333333 0.165644 0.40699 
 Specie.Code.Location.2       time_cts                              0.001633 0.04041 

```{r}
0.200537+0.110581+0.044341+0.015569+0.032118+0.240568+0.165644+0.001633
0.190261+0.151596+0.026476+0.257292+0.196650+0.001756
```


Conditional model:
 Groups                 Name                                  Variance Std.Dev. Corr      
 Specie.Code            (Intercept)                           0.190261 0.43619            
 Specie.Code.session    (Intercept)                           0.151596 0.38935            
 year.session           (Intercept)                           0.026476 0.16272            
 Specie.Code.Location   (Intercept)                           0.257292 0.50724            
 Specie.Code.Location.1 as.ordered(time_cts)0.333333333333333 0.196650 0.44345  0.33 (ar1)
 Specie.Code.Location.2 time_cts                              0.001756 0.04191            


```{r}
fam_nb2 <- "nbinom2"
fit_nb2_corr_FULL_RF <- glmmTMB(form_glmm_ar1_FULL_RF,
                        family = fam_nb2,
                        data = ecuador)
diagnose(fit_nb2_corr_FULL_RF)

fit_nb2_corr_no_sppSess <- update(fit_nb2_corr_FULL_RF, 
                                   . ~ . -(1|Specie.Code:session))
fit_nb2_corr_no_sppSessLoc <- update(fit_nb2_corr_FULL_RF, 
                                   . ~ . -(1|Specie.Code:session:Location))

fit_nb2_corr_no_yrSess <- update(fit_nb2_corr_FULL_RF, 
                                   . ~ . -(1|year:session))

fit_nb2_corr_no_yrSessLoc <- update(fit_nb2_corr_FULL_RF, 
                                   . ~ . -(1|year:session:Location))


bbmle::AICtab(fit_nb2_corr_FULL_RF,
              fit_nb2_corr_no_sppSess,
              fit_nb2_corr_no_sppSessLoc,
              fit_nb2_corr_no_yrSess,    # very good
              fit_nb2_corr_no_yrSessLoc)
```


```{r}
bbmle::AICtab(fit_pois_corr_FULL_RF,
              fit_pois_corr_no_sppSess,
              fit_pois_corr_no_sppSessLoc,
              
              fit_pois_corr_orig,
              
              fit_pois_drop_randslope,
              fit_pois_drop_randslope_from_orig,
              #fit_nb2_corr_FULL_RF,
              fit_nb2_corr_no_sppSess,
              fit_nb2_corr_no_sppSessLoc,
              fit_nb2_corr_no_yrSess#,     # very good
              #fit_nb2_corr_no_yrSessLoc
              )
```



```{r}
model_list <- list(fit_pois_corr_FULL_RF = fit_pois_corr_FULL_RF,
              fit_pois_corr_no_sppSess = fit_pois_corr_no_sppSess,
              fit_pois_corr_no_sppSessLoc = fit_pois_corr_no_sppSessLoc,
              fit_pois_corr_orig = fit_pois_corr_orig,
              fit_pois_drop_randslope = fit_pois_drop_randslope,
              fit_pois_drop_randslope_from_orig = fit_pois_drop_randslope_from_orig,
              
              #fit_nb2_corr_FULL_RF,
              fit_nb2_corr_no_sppSess = fit_nb2_corr_no_sppSess,
              fit_nb2_corr_no_sppSessLoc = fit_nb2_corr_no_sppSessLoc,
              fit_nb2_corr_no_yrSess = fit_nb2_corr_no_yrSess#,    # very good
              #fit_nb2_corr_no_yrSessLoc
              )
```



```{r}
overdisp.stat <- rep(NA,length(model_list))
overdisp.p <- overdisp.stat

for(i in 2:length(list)){
  x <- try(DHARMa::testDispersion(model_list[[i]], type = "DHARMa"), silent = T)
  
  overdisp.stat[i] <- x$statistic
  overdisp.p[i] <- x$p.value
}
```


```{r}
simulationOutput <- DHARMa::simulateResiduals(fittedModel = fit_nb2_corr_no_yrSess, 
                                      refit = F,
                                      plot = F)


plot(simulationOutput)
DHARMa::testDispersion(simulationOutput)
DHARMa::testUniformity(simulationOutput)
DHARMa::testZeroInflation(simulationOutput)
DHARMa::testOutliers(fit_nb2_corr_no_yrSess,
                     type = "bootstrap")


ecuador$resids <- resid(simulationOutput)
```

