---
title: "e) Model community abundance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{e) Model community abundance}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = F
)
```

## Introduction

Model community abundance to test functioning of glmmTMB and determine if there is an overall decline in total captures over time.




```{r setup, warning = F, message = F}
library(avesdemazan)

# Model fitting
library(lme4)       # glmer() for glmm
library(glmmTMB)    # glmm with AR-1  
# library(blme)       # bglmer (used?)

# # Model evaluation
# library(afex)       # functionall_fit()
# library(multcomp)   
# library(emmeans)
# library(bbmle)      # aictab
# library(arm)        # se.fit()

# Result visualization
library(ggplot2)    

#  markdown etc
library(knitr)    
```



## Model TOTAL individuals captures 

Load data.  All individuals captures within a session totaled up, ignoring species. 

```{r}
data(ecuador_tot)
```


The data is 86  by 6
```{r}
dim(ecuador_tot)
```

The data frame looks like this:
```{r}
head(ecuador_tot, 10)
```


```{r}
summary(ecuador_tot)
```



**NOTE**: This shows total captures of species that are *included in the GLMM*, NOT total captures of ALL species. Species that did not have enough captures to meet the inclusion critera are not included, and unbanded individuals are not included.

<!-- TODO: create a datarame with ALL captures regarless of whether they are in the GLMM or not -->


Split data by `session` and plot by `year`.  This is **NOT** how the data are modeled because time is set up as a continuous variable.


```{r, fig.cap = "Within-season changes in total captures.  Declines appear to be occuring in all seasons except SEGUNDA-LLAV"}
ggplot(data = ecuador_tot,
        aes(y = N/tot_net_hours*1000,
            x = year,
            color = session)) +
  geom_point() +
  facet_wrap(~Location) + 
  geom_smooth(method = lm,se = F) + 
  theme_bw()
```

Ignore sessions and plot by `time_cts`.
```{r, fig.cap="Changes in abundance using time as a continous variable"}
ggplot(data = ecuador_tot,
        aes(y = N/tot_net_hours*1000,
            x = time_cts)) +
  geom_point() +
  facet_wrap(~Location) + 
  geom_smooth(method = lm,se = F) + 
  theme_bw() 
```


## Models of community abundance change

### GLM model of community abundance change

Basic GLM for model exploration.

```{r}
glm_fit <- glm(N ~ 1 + time_cts +   # continuous time 
                     Location +     # Location
                     time_cts*Location +  # 
                     offset(log(tot_net_hours)),
                   family = poisson,
                   data = ecuador_tot)
```


Time is significant, but with no time_cts*:Location interaction.
```{r}
summary(glm_fit)
```



### Basic GLMM  with lme4::glmer()

Poisson GLMM; no autocorrelation, poisson errors.

### glmer() Modeling with fixed effects for time*location

```{r}
glmm_fit01 <- lme4::glmer(N ~ 1 + time_cts +   # Null
                          Location +
                          Location*time_cts +
                     #(1|session) +
                     #(1|Location) +
                     (1|year:session) + # all captures w/in a session-year
                     (1|year:session:Location) + # all caps w/in a session-year-loc  
                     #(0+time_cts|Location) + # rand slope for location
                     offset(log(tot_net_hours)),
                   family = poisson,
                   data = ecuador_tot)

summary(glmm_fit01)
```

### glmer() Modeling with ranefs for time*location

Not significant
```{r}
glmm_fit02 <- lme4::glmer(N ~ 1 + time_cts +   # Null
                          #Location +
                          #Location*time_cts +
                     #(1|session) +
                     #(1|Location) +
                     (1|year:session) + # all captures w/in a session-year
                     (1|year:session:Location) + # all caps w/in a session-year-loc  
                     (0+time_cts|Location) + # rand slope for location
                     offset(log(tot_net_hours)),
                   family = poisson,
                   data = ecuador_tot)

summary(glmm_fit02)

# fixef(glmm_fit)[2]
# slopes_loc <- fixef(glmm_fit)[2]+ranef(glmm_fit)$Location
# 
# CI.up
# CI.lo <- fixef(glmm_fit)[2]-ranef(glmm_fit)$Location
# 
# 
# +se.ranef(glmm_fit)$Location*2
# fixef(glmm_fit)[2]-se.ranef(glmm_fit)$Location*2
```



### Poisson glmmTMB without ar1


```{r}
library(splines)
glmmTMB_poisson <- glmmTMB(N ~ 1 + time_cts +   # Null
                          Location +
                          Location*time_cts +
                     (1|year:session) + # all captures w/in a session-year
                     (1|year:session:Location) + # all caps w/in a session-year-loc  
                     (0+time_cts|Location) + # rand slope for location
                     offset(log(tot_net_hours)),
                   family = poisson,
                   data = ecuador_tot)

summary(glmmTMB_poisson)
```




