---
title: "g) Model selection - random effects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{g) Model selection - random effects}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This script assess how well the random effects work.



## Preliminaries

```{r setup}
# Data and functions
library(avesdemazan)

# Model fitting
library(lme4)       # glmer() for glmm
library(glmmTMB)    # glmm with AR-1  
# library(blme)       # bglmer (used?)

# Model evaluation
# library(afex)       # functionall_fit()
# library(multcomp)   
# library(emmeans)
library(bbmle)      # aictab
library(arm)        # se.fit()

# Result visualization
# library(ggplot2)    
# library(ggstance)  
# library(grid)
# library(gridExtra)  

#  markdown etc
library(knitr)    
```



### Load data

Load capture data. 

```{r}
data("ecuador")
```

Key columns for models run in tis script:
```{r}
head(ecuador[,1:9])
```


## Species-specifc models of captures

### Formulas

#### Possible random effects

* There are two general classes of random effects to consider: **location-species**, and **time**
* There's a tradeoff between what deserves to be a random effect, and what will actually converge

**Species-Location:**

* Species are (partially) crossed with location.  
* There are only 3 locations so its difficult to use a random effect just for it.  
* The easiest way to represent this is with a species:location random effect.

**Time random effects:**

* There are two possible random effects for time: sampling year and sampling sessions (~seasons)
* There are only 3 sampling sessions so on their own these work best as a fixed effect, if included at all
* However, we're using a autocorrelated AR-1 model, which I think should impact how we think about the time random effects
* Year as a stand-alone random effect is problematic because it splits up adjacent sampling seasons that occur in different years, eg. 2010-Fall vs. 2011-Winter are adjacent to each other in the sequence of sampling dates but would be grouped into two different bins of the random effect.
* I think what makes the most sense is a `year:sessions` random effect.  With this, all captures at approximately the same time will be induced to be correlated.  Within a nominal year the AR-1 should deal with correlation between successive sampling sessions (Winter-Spring, Spring-Fall).
* It may be useful to include a `session` and `Location*session` FIXED effect to capture the fact there may be typical seasonal dynamics (e.g. altitudinal migration); however, this is likely to be species-specific.
* In theory there might be `Species:session` or even S`pecies:session:Location` effects to account for species-specific variation in seasonal behavior, but this would be hard to deal with and would probably be mostly covered by the `Species:location random` effects and random slopes


## Models

### Model structures

Model structure with NO autocorrelation
```{r}
form_glmm_no_autocorr_FULL_RF <- formula(N ~ 1 +          # Intercept
                      Location +                  # Location FE; 3 level factor
                      # session +                 # session FE
                      # Location*session  +       # Location*session  FE
                      time_cts +                  # Continuous time variable
                      (1|Specie.Code) +         # species RE; DOES THIS CONVERGE?
                      (1|Specie.Code:session) +         #DOES THIS CONVERGE?
                      (1|year:session) +         # DOES THIS CONVERGE? 
                      (1|Specie.Code:Location) +  # Species:location intercept
                      (time_cts + 
                       0|Specie.Code:Location) +   # Species-location RS
                      offset(log(tot_net_hours))   # offset for effort
                      )

```

Model structure WITH autocorrelation
```{r}
# NOTE: additional random effects will make it MUCH more difficult to calculate
# SE!!!!!!!!!!!!!!!!!!!!!!
form_glmm_ar1_FULL_RF <- formula(N ~ 1 +          # Intercept
                      Location +                  # Location FE; 3 level factor
                      # session +                 # session FE
                      # Location*session  +       # Location*session  FE
                      time_cts +                  # Continuous time variable
                      (1|Specie.Code) +           # species RE
                      
                      # session effects    
                      ## species:session; overfitting?  
                      (1|Specie.Code:session) +   # 
                      
                      ## corr within the same sampling period    
                      (1|year:session) +          # 
                        
                      # not necessary b/c dropped in rand slope statement?
                      (1|Specie.Code:Location) +  # Species:location intercept 
                     
                       ar1(as.ordered(time_cts) + 0|Specie.Code:Location) +
                      
                      (time_cts + 
                       0|Specie.Code:Location) +   # Species-location RS
                      
                      offset(log(tot_net_hours))   # offset for effort
                      )
```



Model structure WITH autocorrelation, AND only REs included in ES models:
`(1|Specie.Code:Location) `

This should be (mostly) compatible with all the subsequent code Es wrote, while the model formula above has additional REs which will need to be integrated into code the calculates e.g. standard errors.
```{r}
form_glmm_ar1_ES_RF <- formula(N ~ 1 +          # Intercept
                      Location +                  # Location FE; 3 level factor
                      # session +                 # session FE
                      # Location*session  +       # Location*session  FE
                      time_cts +                  # Continuous time variable
                      #(1|Specie.Code) +           # species RE
                      #(1|Specie.Code:session) +   # 
                      #(1|year:session) +          # 
                      (1|Specie.Code:Location) +  # <===Species:location intercept 
                      ar1(as.ordered(time_cts) + 0|Specie.Code:Location) +
                      
                      (time_cts + 
                       0|Specie.Code:Location) +   # Species-location RS
                      
                      offset(log(tot_net_hours))   # offset for effort
                      )
```



## Fit models

### Poisson models 

As demonstrated previously, Poisson model fit with no computational issues, except a model with only one random intercept (as was run by ES has potential issues with the Wald CI, which according to the diagnose() command should be cleared up by using other ways of formulating CIs..

The code below runs poisson 

* with and without ar1
* with some random effects dropped
* in equivalent forms set up by ES

In order to probe the adequacy of the random effects structure and make sure there are no typos.


```{r}
fam_poisson <- "poisson"

# full error structure, no autocorr
fit_pois_FULL_RF <- glmmTMB(form_glmm_no_autocorr_FULL_RF,         
                      family = fam_poisson,
                      data = ecuador)
diagnose(fit_nb2_FULL_RF)

# full error structu, with autocorr
fit_pois_corr_FULL_RF <- glmmTMB(form_glmm_ar1_FULL_RF,
                        family = fam_poisson,
                        data = ecuador)
diagnose(fit_pois_corr_FULL_RF)


# Drop  (1|Specie.Code)
## No corr, drop (1|Specie.Code)
fit_pois_no_spp_RF <- update(fit_pois_FULL_RF, . ~ . -(1|Specie.Code))
diagnose(fit_pois_no_spp_RF)

## with corr (1|Specie.Code)
fit_pois_corr_no_spp_RF <- update(fit_pois_corr_FULL_RF, . ~ . - (1|Specie.Code))
diagnose(fit_pois_corr_no_spp_RF)


# Drop (1|year:session)
## No corr - Drop (1|year:session)
fit_pois_no_yearSession_RF <- update(fit_pois_FULL_RF, . ~ . - (1|year:session))
diagnose(fit_pois_no_yearSession_RF)

## with Corr,   Drop (1|year:session)
fit_pois_corr_no_yearSession_RF <- update(fit_pois_corr_FULL_RF, . ~ . - (1|year:session))
diagnose(fit_pois_corr_no_yearSession_RF)


# Drop (1|year:Location)
## No corr - Drop (1|year:Location)
fit_pois_no_sppLoc_RF <- update(fit_pois_FULL_RF, . ~ . - (1 | Specie.Code:Location))
diagnose(fit_pois_no_sppLoc_RF)

## Corr -  Drop (1|year:Location)
fit_pois_corr_no_sppLoc_RF <- update(fit_pois_corr_FULL_RF, . ~ . - (1| Specie.Code:Location))
diagnose(fit_pois_corr_no_sppLoc_RF)










# Drop 2 random effects
## no corr - (1|year:session) - (1|Specie.Code) 
fit_pois_no_yrSess_no_spp_RF <- update(fit_pois_FULL_RF, . ~ . - (1|year:session) - (1|Specie.Code) )
diagnose(fit_pois_no_yrSess_no_spp_RF)

## with corr
fit_pois_corr_no_yrSess_no_spp_RF <- update(fit_pois_corr_FULL_RF, . ~ . - (1|year:session) - (1|Specie.Code))
diagnose(fit_pois_corr_no_yrSess_no_spp_RF)




# Drop 3 REs
## Only random intercept in Emily's model was: 
## (1|Specie.Code:Location) 

##
fit_pois_no_yrSess_no_spp_nosppSess <- update(fit_pois_FULL_RF, . ~ . - (1|year:session) 
                                                 - (1|Specie.Code) 
                                                 - (1|Specie.Code:session))
diagnose(fit_pois_no_yrSess_no_spp_nosppSess)

##
fit_pois_corr_no_yrSess_no_spp_nosppSess <- update(fit_pois_corr_FULL_RF, . ~ . - (1|year:session) 
                                                      - (1|Specie.Code) 
                                                      - (1|Specie.Code:session))
diagnose(fit_pois_corr_no_yrSess_no_spp_nosppSess)



# should be same as previous model
fit_pois_corr_ES_RF <- glmmTMB(form_glmm_ar1_ES_RF,
                        family = fam_poisson,
                        data = ecuador)

AIC(fit_pois_corr_no_yrSess_no_spp_nosppSess)
AIC(fit_pois_corr_ES_RF)

```


### Negative binomial models - confirm bad

After updating minor mis-specificaiton of models fit by ES, nb2 models no longer converge or have Hessians etc that are ok.  The code before fits these nb2 models again to confirm this.  This was done in a previous script, but it repeated here to make sure that this is still true with the given the current random effects structure given above.  (If these models stop having issues, the code should be check to make sure the correct models are being run and no changes / typos have entered into the code)  


```{r}
# negative-binomial-2
## Typically the best error structure
fit_nb2_FULL_RF <- glmmTMB(form_glmm_no_autocorr_FULL_RF,         
                      family = nbinom2,#nb2
                      data = ecuador)
diagnose(fit_nb2_FULL_RF)

# throws error
fit_nb2_corr_FULL_RF <- glmmTMB(form_glmm_ar1_FULL_RF,
                        family = nbinom2,
                        data = ecuador)
diagnose(fit_nb2_corr_FULL_RF)

x <- afex:all_fit(fit_nb2_corr_FULL_RF)


# No corr, drop (1|Specie.Code)
## no error
fit_nb2_no_spp_RF <- update(fit_nb2_FULL_RF, . ~ . - (1|Specie.Code))
diagnose(fit_nb2_no_spp_RF)

# Drop  (1|Specie.Code)
## throws convergence warning
fit_nb2_corr_no_spp_RF <- update(fit_nb2_corr_FULL_RF, . ~ . - (1|Specie.Code))
diagnose(fit_nb2_corr_no_spp_RF)

  
  
# No corr - Drop (1|year:session)
fit_nb2_no_yearSession_RF <- update(fit_nb2_FULL_RF, . ~ . - (1|year:session))
diagnose(fit_nb2_no_yearSession_RF)

# Corr -  Drop (1|year:session)
fit_nb2_corr_yearSession_RF <- update(fit_nb2_corr_FULL_RF, . ~ . - (1|year:session))
diagnose(fit_nb2_corr_yearSession_RF)
```


## Emily Scott's original models


### nb models

These models have no issues related to fitting them, but the second one is mis-specified as detailed elsewhere.  They are fit here for comparison purposes to determine how updating the model specification impacts model performance out output.

```{r}
# not autocorr
## I believe there are no structural problems with this
fit_nb2_emily <- glmmTMB(N ~ 1 +                                                    # Null
                Location +                                               # Location
                time_cts +                                               # Continuous time variable
                     (1|Specie.Code:Location) +                               # Species-level intercept'
                     (time_cts + 0|Specie.Code:Location) +
                     offset(log(tot_net_hours)),
                   family = nbinom2,
                   data = ecuador)

diagnose(fit_nb2_emily)

# autocorr
## NOTE - has 
##  ar1(as.ordered(time_cts) + 0|Specie.Code)
## instead of 
##  ar1(as.ordered(time_cts) + 0|Specie.Code:Location)
fit_nb2_corr_emily <- glmmTMB(N ~ 1 +                                           # Null
                          Location +                                               # Location
                          time_cts +                                       # Continuous time variable
                          (1|Specie.Code:Location) +                       # Species-level intercept
                          (time_cts + 0|Specie.Code:Location) +
                          ar1(as.ordered(time_cts) + 0|Specie.Code) + # AR across time
                          offset(log(tot_net_hours)),
                        family = nbinom2,
                        data = ecuador)
diagnose(fit_nb2_corr_emily)
```

### poison models




```{r}
# not autocorr
## I believe there are no structural problems with this
fit_pois_emily <- glmmTMB(N ~ 1 +                                                    # Null
                     Location +                                               # Location
                     time_cts +                                               # Continuous time variable
                     (1|Specie.Code:Location) +                               # Species-level intercept'
                     (time_cts + 0|Specie.Code:Location) +
                     offset(log(tot_net_hours)),
                   family = poisson,
                   data = ecuador)

diagnose(fit_pois_emily)

# autocorr
## NOTE - has 
##  ar1(as.ordered(time_cts) + 0|Specie.Code)
## instead of 
##  ar1(as.ordered(time_cts) + 0|Specie.Code:Location)
fit_pois_corr_emily <- glmmTMB(N ~ 1 +                                           # Null
                          Location +                                               # Location
                          time_cts +                                       # Continuous time variable
                          (1|Specie.Code:Location) +                       # Species-level intercept
                          (time_cts + 0|Specie.Code:Location) +
                          ar1(as.ordered(time_cts) + 0|Specie.Code) + # AR across time
                          offset(log(tot_net_hours)),
                        family = poisson,
                        data = ecuador)
diagnose(fit_pois_corr_emily)

# update
## NOTE: throws error regarding Walk CI
fit_pois_corr_emily_fixed <- glmmTMB(N ~ 1 +                                           # Null
                          Location +                                               # Location
                          time_cts +                                       # Continuous time variable
                          (1|Specie.Code:Location) +                       # Species-level intercept
                          (time_cts + 0|Specie.Code:Location) +
                          ar1(as.ordered(time_cts) + 0|Specie.Code:Location) + # Autocorrelation across time
                          offset(log(tot_net_hours)),
                        family = poisson,
                        data = ecuador)
diagnose(fit_pois_corr_emily_fixed)

# this should be the same as other models ; this was run just to confirm 
# my understanding of the model structure / confirm no typos
AIC(fit_pois_corr_no_yrSess_no_spp_RF_nosppSess)
AIC(fit_pois_corr_ES_RF)
AIC(fit_pois_corr_emily_fixed)
```








## Output 

### Compare all models

Models fit by ES have high AIC.

nb2 models are generally (always?) worse than corresponding poisson

```{r}
bbmle::AICctab(fit_nb2_emily,
               fit_nb2_corr_emily,
  
               fit_nb2_no_spp_RF,
               fit_nb2_corr_FULL_RF,   # was best, but mis-specified
               fit_nb2_corr_no_spp_RF,
               fit_nb2_no_yearSession_RF,
               fit_nb2_corr_yearSession_RF,
               fit_nb2_FULL_RF,
               fit_pois_FULL_RF ,
fit_pois_corr_FULL_RF ,
fit_pois_no_spp_RF ,
fit_pois_corr_no_spp_RF, 
fit_pois_no_yearSession_RF, 
fit_pois_corr_no_yearSession_RF ,
fit_pois_no_sppLoc_RF ,
fit_pois_corr_no_sppLoc_RF, 
fit_pois_no_yrSess_no_spp_RF, 
fit_pois_corr_no_yrSess_no_spp_RF,
fit_pois_no_yrSess_no_spp_RF_nosppSess,
fit_pois_corr_no_yrSess_no_spp_RF_nosppSess, 
fit_pois_corr_ES_RF )
```


### Poisson models

Full random effects structure works best

```{r}
bbmle::AICctab(fit_pois_corr_FULL_RF ,
fit_pois_no_spp_RF ,
fit_pois_corr_no_spp_RF, 
fit_pois_no_yearSession_RF, 
fit_pois_corr_no_yearSession_RF ,
fit_pois_no_sppLoc_RF ,
fit_pois_corr_no_sppLoc_RF, 
fit_pois_no_yrSess_no_spp_RF, 
fit_pois_corr_no_yrSess_no_spp_RF,
fit_pois_no_yrSess_no_spp_RF_nosppSess,
fit_pois_corr_no_yrSess_no_spp_RF_nosppSess, 
fit_pois_corr_ES_RF
)
```



## Save the formula objects

Moving forward I'll work with two models:
 
* full random effects
* just spp-location fiexed effects

The spp-location fixed effect model is the same structure worked with by ES,
just with Poisson error and updated ar1() structure.  This model will be (mostly?)
compatible with current code; the more complex model will require that code for calculating standard errors etc be updated.

```{r}
# Full model
usethis::use_data(fit_pois_corr_FULL_RF, overwrite = TRUE)

# Reduced model
AIC(fit_pois_corr_ES_RF, 
    fit_pois_corr_no_yrSess_no_spp_RF_nosppSess)

fit_pois_corr_only_sppLoc <- fit_pois_corr_no_yrSess_no_spp_RF_nosppSess
usethis::use_data(fit_pois_corr_only_sppLoc, overwrite = TRUE)

# other models for comparison
AIC(fit_nb2_corr_emily, 
    fit_pois_corr_emily)
fit_nb2_corr_misspec  <- fit_nb2_corr_emily
fit_pois_corr_misspec <-  fit_pois_corr_emily


usethis::use_data(fit_nb2_corr_misspec, overwrite = TRUE)
usethis::use_data(fit_pois_corr_misspec, overwrite = TRUE)


```


