---
title: "h) Process best model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{h) Process best model}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(avesdemazan)
```

## Introduction

This script

1. Re-fits the best model
1. Extracts the random slopes




## Preliminaries



### Load packages
```{r House-keeping, warning = FALSE, message = F, echo = T, eval = TRUE}
# Data cleaning
library(dplyr)      
library(tidyr)   
library(reshape2)
library(stringi)  


# Model fitting
library(lme4)       # glmer() for glmm
library(glmmTMB)    # glmm with AR-1  
library(blme)       # bglmer (used?)

# Model evaluation
library(afex)       # functionall_fit()
library(multcomp)   
library(emmeans)
library(arm)        # se.fit()

# Result visualization
library(ggplot2)    
library(ggstance)  
library(grid)
library(gridExtra)  

#  markdown etc
library(knitr)     

  
```


afex::all_fit 


## Best model 


```{r Show best model specification, eval = F, include = TRUE, warning = FALSE}
# This is the best model structure
data(fit_pois_corr_FULL_RF)

# This is a reduced model structure
data(fit_pois_corr_only_sppLoc)

# mis-specified models, for comparison
data(fit_nb2_corr_misspec)  # nb mis-spec
data(fit_pois_corr_misspec) # pois mis-spec

# check modesl
diagnose(fit_pois_corr_FULL_RF)
diagnose(fit_pois_corr_only_sppLoc) # has issue with Wald stats
diagnose(fit_nb2_corr_misspec)
diagnose(fit_pois_corr_misspec) # has issue with Wald stats
```


Save best model to the object `best_model`.
```{r}
best_model <- fit_pois_corr_FULL_RF
usethis::use_data(best_model,
                  overwrite = TRUE)
usethis::use_r("best_model")
```



## Examine best model

Examine best model and compare with ES-type model

```{r}

fixef(best_model)
fixef(fit_pois_corr_only_sppLoc)

AIC(best_model,fit_pois_corr_only_sppLoc)

```


```{r}
summary(best_model)
summary(fit_pois_corr_only_sppLoc)
```



## Look at random effects

What is the model formula?
```{r}
formula(best_model)
```


```{r}
x <- summary(best_model)
x$varcor
```

Get the species-specific random slopes, which are nested within Location

random-slopes specification
              (time_cts + 0 | Specie.Code:Location)

ar1 specification:
ar1(as.ordered(time_cts) +  0 | Specie.Code:Location)
```{r}
# best model
ranefsx <- ranef(best_model)

# reduxed model
ranefsx_redox <- ranef(fit_pois_corr_only_sppLoc) #reduced model

# mis-specified model
ranefsx_nb_misspec <- ranef(fit_nb2_corr_misspec)
ranefsx_pois_misspec  <- ranef(fit_pois_corr_misspec)



# overall rand eff object
names(ranefsx)

# zi = zero inflated; empty
ranefsx$zi

# location of what we want
## random slopes in "Specie.Code:Location" 
## AR1 stuff     also in"Specie.Code:Location" 
names(ranefsx$cond)

# everything associate with Specie.Code:Location
## intercepts
## rand slopes
## ar1

# best model
sppLocRE    <- ranefsx$cond$`Specie.Code:Location`

# for comparison
sppLocRE_redox <- ranefsx_redox$cond$`Specie.Code:Location`
sppLocRE_poismisSpec    <- ranefsx_nb_misspec$cond$`Specie.Code:Location`
sppLocRE_nbmisSpec     <- ranefsx_pois_misspec$cond$`Specie.Code:Location`


dim(sppLocRE) == dim(sppLocRE_ES)

plot(sppLocRE$time_cts ~ sppLocRE_redox$time_cts, main = "Best vs. redopx")
abline(a = 0, b = 1)
plot(sppLocRE$time_cts ~ sppLocRE_poismisSpec$time_cts, 
     main = "Best vs mis-spec pois")
abline(a = 0, b = 1)
plot(sppLocRE$time_cts ~ sppLocRE_nbmisSpec$time_cts, 
     main = "Best vs mis-spec nb")
abline(a = 0, b = 1)

abline(a = 0, b = 1)

# get number of species
head(ranefsx$cond$`Specie.Code:Location`)
N_randslopes <- dim(ranefsx$cond$`Specie.Code:Location`)[1] # 79 x 33
N_randslopes


```

Snapshot of the random slope

**NOTE: these random slopes are deviations from the OVERALL slope parameter of the model!**
```{r}
head(ranefsx$cond$`Specie.Code:Location`)
```


```{r}
vcov(best_model)
```




## Process random slopes


Extract SD / var using TMB::sdreport()

"After optimization of an AD model, sdreport() is used to calculate standard deviations of all model parameters"

* `getJointPrecision	= Optional`. Return full joint precision matrix of random effects and parameters?
  + From `sdreport` doc:"The full joint covariance is not returned by default, because it may require large amounts of memory. It may be obtained by specifying `getJointPrecision=TRUE`, ..."
* `bias.correct`	= logical indicating if bias correction should be applied


### Get TMB sd report

This takes a second
```{r}
s1 <- TMB::sdreport(obj = best_model$obj, 
                    getJointPrecision = TRUE, 
                    bias.correct = TRUE)
```

The summary output of `sdreport` are the coefficients and SE for the **main effects** of the model
```{r}
s1
```

This is the same result
```{r}
summary(s1,"fixed")
```

This is basically the sames as `summary()` on the original model object
```{r}
summary(s1,"fixed", p.value = TRUE)
```


The summary for the random effects
```{r}
head(summary(s1, "random") , 10)

raneff_summary <- as.data.frame(summary(s1, "random"))
```

There are **MANY** random effects coefficients 
```{r}
dim(summary(s1, "random"))
```


The output of the `sdreport()` is complex
```{r}
is(s1)
names(s1)

# covariance matrix of fixed effects
s1$cov.fixed

```

The diagonal of the SE cov matrix is 1336 
(OLD comment by NLB: "I assume the whole matrix isn't available b/c its big and/or b/c its not very useful"; not sure what I meant by this....)
```{r}
x <- s1$diag.cov.random
x <- as.data.frame(x)
dim(x) # 1336


```


Compare the summary output to the covariance matrix diagonal
```{r}
raneff_summary2 <- cbind(raneff_summary, x)
dim(raneff_summary2)
head(raneff_summary2)
```


### Convert 

```{r}
s2 <- sqrt(s1$diag.cov.random)
length(s2) #1336

# ? 
## 1345 x 1345
test <- solve(s1$jointPrecision)

1345-1336

```

Extract something to get SD
```{r}
#?
## output is just 0 (?)
parameters <- best_model$obj$env$par
parameters <- as.data.frame(parameters)
head(parameters) 

# get parameters names
## 4 "beta", rest "b"
## 4 betas = intercept, time_cts, etc
parameters$names <- names(best_model$obj$env$par)
head(parameters)

# get sd from "test" object 
parameters$sd <- sqrt(diag(test))

dim(parameters)
```


Create column for name of coefficient

NOTE: here we have hard-coded indices

```{r}
i <- N_randslopes
length(parameters$variable) # 1345
parameters$variable <- NA  # create new column
parameters$variable[1:4] <- c("(Intercept)","LocationMAIN","LocationMASE","time_cts")

parameters$variable[5:162] <-   c(rep("Intercept", i), 
                                  rep("time_cts", i))
# only 162 parameters
## but parameters objects is 1345; what are the rest?
4 + N_randslopes*2

1345 - c(4 + N_randslopes*2) # 1183

# There are a LOT of parameters associated with the AR1 stuff
## about 38 spp x 31 time points
38*31

# This accounts for most of the rows in parameters; 
## 1345: size of joint precisions matrix
## 4 = number of fixed effects
## 2*N_randslopes b/c intercept and slope
## AR1 stuff is 38 * 31
1345 - c(4 + 2*N_randslopes) - 38*31
```




Name coefficients 
```{r}
parameters$spp <- NA
fixed_eff_names <- rep("Population", 4) #why population?
rand_eff_intercept_names <- rownames(ranef(best_model)[[1]]$`Specie.Code:Location`)
rand_eff_slope_names <- rand_eff_intercept_names
parameters$spp[1:162] <- c(fixed_eff_names, 
                           rand_eff_intercept_names,    
                           rand_eff_slope_names)
```


loop over x and combine y into sd

TODO: check - things are hard coded here - does that matter?

I think ES codded the delta method here

```{r}
# slope estimate = -0.028207, SE = 0.009887
test[4,4] # SD for slope
sqrt(test[4,4]) 


```

```{r}
parameters$aggregate_sd <- NA

# Slopes start at 84
head(parameters[75:88,])
N_randslopes+4 # last intercept
N_randslopes+4 +1  #first slope

slope.i <- 4

# note: i starts at 0
for (i in 0:78)
{ 
  
  # combined SD                  
  parameters$aggregate_sd[84+i] <- sqrt(test[slope.i,slope.i] + # fixef slope var
                                                test[84+i,84+i] +     # ranefslope var  
                                                2*test[84+i,slope.i]) # fixef-ranef covariance
}

```



## Extract SE 

TODO: check - things are hard coded here - does that matter?

```{r Store standard errors, echo = T, fig.height = 7.25}
N_randslopes+4 + 1  #first slope
2*N_randslopes + 4   #last slope

# get subst that we want
subset_for_merge <- parameters[84:162,c("spp","aggregate_sd")]
colnames(subset_for_merge)[1] <- "id"

sd_int   <- s2[1:79]
sd_slope <- s2[80:158]

slopes <- ranef(best_model)[[1]]$`Specie.Code:Location` # 79 x 2
dim(slopes)
#ranef(best_model)[[1]]$`Specie.Code` # 38 x 33
# 79 * 2 + 38 * 33

s0 <- best_model$sdr

                       #time coefficient                      # ? individual time slope?
slopes$time_cts_tot <- summary(best_model)[[6]]$cond[4,1] + slopes$time_cts
slopes$time_sd  <- sd_slope
slopes$time_lb  <- slopes$time_cts_tot - 1.96*slopes$time_sd
slopes$time_ub  <- slopes$time_cts_tot + 1.96*slopes$time_sd

slopes$spp <- substring(rownames(slopes), 1, 4)
slopes$loc <- substring(rownames(slopes), 6, 9)
slopes$id <- rownames(slopes)

slopes.unique <- slopes[!duplicated(slopes$spp),]

slopes$spp2 <- ordered(slopes$spp, levels = slopes.unique[order(slopes.unique$time_cts_tot),]$spp)
slopes$id2 <- ordered(slopes$id, levels = slopes[order(slopes$time_cts_tot),]$id)

slopes <- left_join(slopes, subset_for_merge, by = "id")

# calculate lower (lb) and upper (ub) bound of CIs
## NOTE: use slopes$time_cts_tot (as shown and as was in orig script)
###      NOTE time_cts
slopes$time_lb2  <- slopes$time_cts_tot - 1.96*slopes$aggregate_sd
slopes$time_ub2  <- slopes$time_cts_tot + 1.96*slopes$aggregate_sd

# ggplot(data = slopes, aes(x = time_cts, y = slopes$spp2)) +
#   geom_point(aes(x = time_cts, y = spp2)) +
#   geom_segment(aes(x = time_lb, xend = time_ub, y = spp2, yend = spp2)) +
#   facet_grid(~loc) +
#   geom_vline(aes(xintercept = 0, color = "No Change")) +
#   geom_vline(aes(xintercept = summary(best_model)[[6]]$cond[4,1], color = "Population-level estimate")) +
#   #geom_rect(aes(xmin = -0.016779 - 1.96*0.003354, xmax = -0.016779 + 1.96*0.003354, ymin = -Inf, ymax = Inf), fill = "blue", alpha = 0.2) +
#   xlab("Species-level time coefficient \n") + ylab("Species Code") +
#   #scale_colour_manual(values = c("Population-level estimate" = "blue")) +
#   scale_colour_manual(values =c("No Change" = "red", "Population-level estimate" = "blue")) +
#   #scale_fill_manual(values = c("blue" = "blue")) +
#   theme(panel.spacing = unit(1, "lines")) +
#   theme(legend.position = "bottom", legend.title = element_blank(), legend.background = element_blank(), legend.box.background = element_rect(colour = "black"))

# habita labels for plots
## main labels

# habitat_labels <- c("Secondary Forest\n(LLAV)", "Primary Forest\n(MASE)", "Introduced Forest\n(MAIN)")
#habitat_labels <- c("Montane shrubland\n(SHRUB)", "Native Forest\n(NATIVE)", "Mixed Native & Introduced Forest\n(MIXED)")
habitat_labels <- c("SHRUB", "NATIVE", "MIXED")

## vector names
# names(habitat_labels) <- c("LLAV","MASE", "MAIN")
names(habitat_labels) <- c("SHRUB","NATIVE", "MIXED")


##
## Species list
data("Ecuador_Species_List")
spp_trt <- Ecuador_Species_List


spp_trt$Species_SciName <- gsub("[\\(\\)]", "", regmatches(spp_trt$Species, gregexpr("\\(.*?\\)", spp_trt$Species)))


spp_trt$Species_SciName <- paste0(substr(spp_trt$Species_SciName,1,1),". ", sub("^\\S+\\s+", '', spp_trt$Species_SciName))


i.col <- which(colnames(spp_trt) %in% c("Specie.Code", "Species_SciName"))

spp_trt[,i.col[1]] <- as.character(spp_trt[,i.col[1]])

slopes.unique <- dplyr::left_join(slopes.unique, spp_trt[,i.col], by = c("spp" = "Specie.Code"))

slopes <- dplyr::left_join(slopes, spp_trt[,i.col], by = c("spp" = "Specie.Code"))

slopes$Species_SciName2 <- ordered(slopes$Species_SciName, levels = slopes.unique[order(slopes.unique$time_cts),]$Species_SciName)

slopes$Sig <- ifelse(slopes$time_ub2 < 0, "Significant", "Not Significant")


```




Site descriptions from MS
"We placed three sampling sites in areas with unique habitat types. 
These included: 
(1) native, mature secondary, subtropical moist broadleaf forest (NATIVE) located in Mazán; 
(2) mixed native and non-native forest (MIXED) also located in Mazán; and
(3) native montane shrubland (SHRUB) located in Llaviuco Valley."


Change labels from those used in earlier drafts of MS 
```{r}
slopes$loc.orig <- slopes$loc

slopes$loc <- gsub("LLAV","SHRUB",slopes$loc)
slopes$loc <- gsub("MAIN","MIXED",slopes$loc)
slopes$loc <- gsub("MASE","NATIVE",slopes$loc)

# set order
slopes$loc <- factor(slopes$loc, levels = c("NATIVE","MIXED","SHRUB"))
slopes$loc <- factor(slopes$loc, levels = c("SHRUB","MIXED","NATIVE"))
```

```{r}
summary(slopes$loc)
```



Save slopes

```{r}
write.csv(slopes, file = "species_specific_slops.csv")
```


```{r}
slopes
```
