---
title: "m) Trait analysis - body size"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{m) Trait analysis - body size}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(avesdemazan)
```


```{r body size fit, echo = T}
BodySize_table <- ecuador_id %>%
  group_by(Band.Number, Body_Size) %>%
  summarize(BodySize_group = length(Body_Size)) %>%
  group_by(Body_Size) %>%
  summarize(N = length(Body_Size))

kable(BodySize_table, caption = "Number of individuals by body size")

ecuador$bodysize_group <- ifelse(ecuador$Body_Size == 1, "BodySize1",
                                 ifelse(ecuador$Body_Size == 2, "BodySize2",
                                        "BodySize3+"))

fit_nb2_corr_bodysize <- glmmTMB(N ~ 1 +                                               # Null
                                  bodysize_group +
                                  #bodysize_group*time_cts +
                                  Location +                                          # Location
                                  time_cts +                                          # Continuous time variable
                                  (1|Specie.Code:Location) +                          # Species-level intercept
                                  (time_cts + 0|Specie.Code:Location) +
                                  ar1(as.ordered(time_cts) + 0|Specie.Code) +         # Autocorrelation across time
                                  offset(log(tot_net_hours)),
                                family = nbinom2,
                                data = ecuador)

BodySize_tab_fit <- summary(fit_nb2_corr_bodysize)$coefficients$cond
rownames(BodySize_tab_fit) <- c("Intercept", "Body Size 2", "Body Size 3+", "Location (MAIN)", "Location (MASE)", "Time")

#kable(BodySize_tab_fit, caption = "Body size model fit on original scale")

# Transformation
vcov <- matrix(unlist(vcov(fit_nb2_corr_bodysize)), nrow = 6)

# LLAV
BodySize_tab_fit2 <- data.frame(Estimate_Llav = c(BodySize_tab_fit[1,1],
                                                  BodySize_tab_fit[1,1] + BodySize_tab_fit[2,1],
                                                  BodySize_tab_fit[1,1] + BodySize_tab_fit[3,1]))

BodySize_tab_fit2$SE_Llav <- c(round(sqrt(vcov[1,1]), 3),
                           round(sqrt(vcov[1,1] + vcov[2,2] + 2*vcov[1,2]), 3),
                           round(sqrt(vcov[1,1] + vcov[3,3] + 2*vcov[1,3]), 3))

BodySize_tab_fit2$LB_Llav <- round(exp(BodySize_tab_fit2$Estimate_Llav - 1.96*BodySize_tab_fit2$SE_Llav), 5)*1000
BodySize_tab_fit2$Est_Llav <- round(exp(BodySize_tab_fit2$Estimate_Llav), 5)*1000
BodySize_tab_fit2$UB_Llav <- round(exp(BodySize_tab_fit2$Estimate_Llav + 1.96*BodySize_tab_fit2$SE_Llav), 5)*1000

BodySize_tab_fit2$CI_Llav <- paste0("(", BodySize_tab_fit2$LB_Llav, ", ", BodySize_tab_fit2$Est_Llav, ", ", BodySize_tab_fit2$UB_Llav, ")")

# MAIN
BodySize_tab_fit2$Estimate_Main <- c(BodySize_tab_fit[1,1] + BodySize_tab_fit[4,1],
                                 BodySize_tab_fit[1,1] + BodySize_tab_fit[2,1] + BodySize_tab_fit[4,1],
                                 BodySize_tab_fit[1,1] + BodySize_tab_fit[3,1] + BodySize_tab_fit[4,1])

BodySize_tab_fit2$SE_Main <- c(round(sqrt(vcov[1,1] + vcov[4,4] + 2*vcov[1,4]), 3),
                           round(sqrt(vcov[1,1] + vcov[2,2] + 2*vcov[1,2] + vcov[4,4] + 2*vcov[1,4] + 2*vcov[2,4]), 3),
                           round(sqrt(vcov[1,1] + vcov[3,3] + 2*vcov[1,3] + vcov[4,4] + 2*vcov[1,4] + 2*vcov[3,4]), 3))

BodySize_tab_fit2$LB_Main <- round(exp(BodySize_tab_fit2$Estimate_Main - 1.96*BodySize_tab_fit2$SE_Main), 5)*1000
BodySize_tab_fit2$Est_Main <- round(exp(BodySize_tab_fit2$Estimate_Main), 5)*1000
BodySize_tab_fit2$UB_Main <- round(exp(BodySize_tab_fit2$Estimate_Main + 1.96*BodySize_tab_fit2$SE_Main), 5)*1000

BodySize_tab_fit2$CI_Main <- paste0("(", BodySize_tab_fit2$LB_Main, ", ", BodySize_tab_fit2$Est_Main, ", ", BodySize_tab_fit2$UB_Main, ")")

# MASE
BodySize_tab_fit2$Estimate_Mase <- c(BodySize_tab_fit[1,1] + BodySize_tab_fit[5,1],
                                 BodySize_tab_fit[1,1] + BodySize_tab_fit[2,1] + BodySize_tab_fit[5,1],
                                 BodySize_tab_fit[1,1] + BodySize_tab_fit[3,1] + BodySize_tab_fit[5,1])

BodySize_tab_fit2$SE_Mase <- c(round(sqrt(vcov[1,1] + vcov[5,5] + 2*vcov[1,5]), 3),
                           round(sqrt(vcov[1,1] + vcov[2,2] + 2*vcov[1,2] + vcov[5,5] + 2*vcov[1,5] + 2*vcov[2,5]), 3),
                           round(sqrt(vcov[1,1] + vcov[3,3] + 2*vcov[1,3] + vcov[5,5] + 2*vcov[1,5] + 2*vcov[3,5]), 3))

BodySize_tab_fit2$LB_Mase <- round(exp(BodySize_tab_fit2$Estimate_Mase - 1.96*BodySize_tab_fit2$SE_Mase), 5)*1000
BodySize_tab_fit2$Est_Mase <- round(exp(BodySize_tab_fit2$Estimate_Mase), 5)*1000
BodySize_tab_fit2$UB_Mase <- round(exp(BodySize_tab_fit2$Estimate_Mase + 1.96*BodySize_tab_fit2$SE_Mase), 5)*1000

BodySize_tab_fit2$CI_Mase <- paste0("(", BodySize_tab_fit2$LB_Mase, ", ", BodySize_tab_fit2$Est_Mase, ", ", BodySize_tab_fit2$UB_Mase, ")")

BodySize_tab_fit2$BodySize <- c("1", "2", "3+")

plot_bodysize <- data.frame(Estimate = c(BodySize_tab_fit2$Est_Llav, BodySize_tab_fit2$Est_Main, BodySize_tab_fit2$Est_Mase),
                        LB = c(BodySize_tab_fit2$LB_Llav, BodySize_tab_fit2$LB_Main, BodySize_tab_fit2$LB_Mase),
                        UB = c(BodySize_tab_fit2$UB_Llav, BodySize_tab_fit2$UB_Main, BodySize_tab_fit2$UB_Mase),                         BodySize = rep(BodySize_tab_fit2$BodySize, 3),
                        Habitat = c(rep("Secondary Forest (LLAV)", 3), 
                                    rep("Introduced Forest (MAIN)", 3),
                                    rep("Primary Forest (MASE)", 3)),
                        Y1 = c(3.15, 3, 2.85,
                              2.15, 2, 1.85,
                              1.15, 1, 0.85),
                        Y2 = c(0.85, 1.85, 2.85,
                               1.15, 2.15, 3.15,
                               1, 2, 3))

plot_bodysize2 <- plot_bodysize[,c(5,4,1,2,3)]
colnames(plot_bodysize2)[4:5] <- c("Lower 95% Estimate", "Upper 95% Estimate")
kable(plot_bodysize2, caption = "Body size model baseline capture rates per 1000 net hours")
```

We find that body size 2 has significantly lower baseline capture rates relative to body size 1. Body size 3 has significantly lower baseline capture rates relative to body size 1. There is no statistically significant difference in how the capture rates change with time for the three body size groups (i.e. all are decreasing by 2.8\% (95\% CI: 0.9-4.7\%) per year from their respective baseline capture rate). There are no statistically significant differences between the habitat types (LLAV, MAIN, MASE) in terms of baseline capture rates.



```{r, echo = T, fig.height= 4}
ggplot(aes(x = Estimate, y = Y1, color = BodySize), data = plot_bodysize) +
  geom_point(size = 1.5) +
  geom_segment(aes(x = LB, xend = UB, y = Y1, yend = Y1), size = 1) +
  theme_bw() +
  scale_y_continuous(breaks = 1:3, labels = c("Primary Forest (MASE)", "Introduced Forest (MAIN)", "Secondary Forest (LLAV)")) +
  ylab("Habitat\n") +
  xlab("Predicted captures per 1000 net hours") +
  ggtitle("Predicted capture rates by body size")

ggplot(aes(x = Estimate, y = Y2, color = Habitat), data = plot_bodysize) +
  geom_point(size = 1.5) +
  geom_segment(aes(x = LB, xend = UB, y = Y2, yend = Y2), size = 1) +
  theme_bw() +
  scale_y_continuous(breaks = 1:3, labels = c("1", "2", "3+")) +
  ylab("Body size\n") +
  xlab("Predicted captures per 1000 net hours") +
  ggtitle("Predicted capture rates by body size")
```

